literature coding explore

### 2024-04-05 reports/lit_codes_explore.qmd
update figures for quarto doc

library(tidyverse)
library(tidyquant)
library(here)

codes_vars <- read.csv(here("data/code_w_vars.csv"))
code_freq <- read.csv(here("data/code_freq2.csv"))
keywords <- read.csv(here("data/keywords.csv"))
occur_all <- read.csv(here("data/overlap_all.csv"))
occur_br <- read.csv(here("data/overlap_bridge.csv"))
occur_ba <- read.csv(here("data/overlap_barrier.csv"))
```

@@

In June 2022 a systematic search was completed using Web of Science to identify articles that mentioned factors that either contributed to "successful" implementation of dynamic ocean management (DOM), hereafter referred to as a "bridge". Literature that discussed factors that would need to be addressed or considered prior to effective implementation of dynamic management approaches or tools were also included, hereafter referred to as a "barrier". In this systematic review, we coded bridges and barriers that were related to DOM voluntary programs or enforced regulations, as well as those that were related to tools (e.g., species spatial modeling tools such as EcoCast) that support dynamic management approaches.

@@

A search was completed in Web of Science for each combination of primary and secondary terms. Works from the primary (i.e., peer reviewed journals) and grey (e.g., theses, book chapters) were included, and this initial search resulted in 1,351 articles. The titles and abstracts for each of these documents were scanned for relevance, and the inclusion of the terms 'dynamic' or 'adaptive' management were confirmed. From this cut, a remaining 127 articles were saved. An additional search through the NOAA Institutional Repository and Federal Register was also completed. Here, documents such as policy reports and planning documents were selected. This search resulted in 795 documents. The titles and abstracts for each of these documents were scanned for relevance, and ultimately, 10 articles were saved for a full read through to determine final inclusion. Inclusion and exclusion criteria for both queries are described here:

@@

Two reviewers went through each of the 110 articles to identify pieces of text that were related to the "successful" implementation of DOM or its recommendation, as well as the barriers or obstacles that hindered the efficacy of this approach or must be considered prior to it's broader application. Through this effort, 17 "bridges" and 9 "barriers" themes were coded. For both the bridges and barriers code categories, three larger groups were established: species, social, and data factors. A draft of the full list of codes/themes are depicted here. Please note that this was a draft of codes, and thus a few themes may be missing or the exact language may be revised over time.

@@

![](images/sunburst_success.png){fig-align="center" width="100%"}

@@

![](images/sunburst_limits.png){width="100%"}
:::

## Visualizing the bridges and barriers themes from the literature

@@

### Code counts/frequencies

>>>>

library(tidyverse)
library(tidyquant)
library(here)
library(kableExtra)
library(waffle)
library(extrafont)
library(hrbrthemes)
library(patchwork)

code_freq <- read.csv(here("data/code_freq_032224.csv"))
codes_vars <- read.csv(here("data/code_w_vars_032224.csv"))
keywords <- read.csv(here("data/keywords_032224.csv"))
occur_all <- read.csv(here("data/code_configs_all_032224.csv"))
occur_br <- read.csv(here("data/code_configs_bridges_032224.csv"))
occur_ba <- read.csv(here("data/code_configs_barriers_032224.csv"))

#change codes to new names 
  #code_freq
    #sp br  
code_freq$code_cat <- replace(code_freq$code_cat, code_freq$code_cat == "Species bridge", "Ecological bridge")
code_freq$code_cat <- replace(code_freq$code_cat, code_freq$code_cat == "Species barrier", "Ecological barrier")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Limited Prey Resources", "Target species is a prey specialist")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Fine Scale management required", "Fine-scale management appropriate for target species")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Migratory and/or dynamic range", "Highly mobile target species")
code_freq$code_name[code_freq$code_cat == "Ecological bridge" & code_freq$code_name == "Climate change"] <- "Expected climate change induced range shifts"
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Known Species Nat History", "Life history information available for target species")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Distinct Habitat between target and non-target species", "Distinct habitat preferences among species in system")

    #data br
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Advances in Tech", "Access to advanced technology and techniques")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Scale appropriate for system", "Management scale aligns with system needs")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Open-Access Public Data", "Data transparency and accessibility")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Availability for species", "Target species data availability")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Stakeholder/Anthropogenic Data Access", "Resource use data availability")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Enviro Access", "Abiotic data availability")

    #soc br
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Open Communication or Co-Mngmt in place", "Co-management and communication")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Adoption Interests & Incentives", "Adoption incentives for end users")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Pre-existing regulation & management institutions", "Pre-existing regulation and management institutions")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "DOM implementation feasible (Socio-econ)", "Feasibility")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Resource Availability (economical)", "Resource availability")

  #sp ba 
code_freq$code_cat[code_freq$code_name == "Model Abuse"] <- "Data barrier"
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Switch to multi-species mngmnt", "Need for multispecies management")

  #data ba
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Gaps and Accuracy", "Gaps and inaccuracy")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Model Abuse", "Model abuse")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Mismatch in scale", "Management scales do not align with system needs")
code_freq$code_cat[code_freq$code_name == "Climate change"] <- "Data barrier"
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Climate change", "Climate change adding uncertainty to target species habitat use")

  #soc ba
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Unintended consequences", "Indirect socioeconomic consequences")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Political permanence", "Lack of political administration permanence")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Communication & Distribution", "Poor communication and outreach")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Recipient participation", "Poor program implementation")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Cost", "Implementation costs")

```

@@

In June 2022 a systematic search was completed using Web of Science to identify articles that mentioned factors that either contributed to "successful" implementation of dynamic ocean management (DOM), hereafter referred to as a "bridge". Literature that discussed factors that would need to be addressed or considered prior to effective implementation of dynamic management approaches or tools were also included, hereafter referred to as a "barrier". In this systematic review, we coded bridges and barriers that were related to DOM voluntary programs and/or enforced regulations, as well as those that were related to tool development (e.g., species spatial modeling tools such as EcoCast) that support dynamic management approaches.

@@

A search was completed in Web of Science for each combination of primary and secondary terms. Works from the primary (i.e., peer reviewed journals) and grey (e.g., theses, book chapters) were included, and this initial search resulted in 1,351 articles. The abstracts for each of these documents were scanned for relevance, and the inclusion of the terms 'dynamic' or 'adaptive' management were confirmed. From this cut, a remaining 127 articles were saved. An additional search through the NOAA Institutional Repository and Federal Register was also completed. Here, documents such as policy reports and planning documents were selected. This search resulted in 795 documents. The abstracts or content summaries for each of these documents were scanned for relevance, and 10 articles were saved for a full read through to determine final inclusion. Inclusion and exclusion criteria for both queries are described here:

@@

Two reviewers went through each of the 110 articles to identify pieces of text that were related to DOM bridges that were linked to the approaches success or were used as justification when systems recommended its adoption. The reviewers also identified the barriers that hindered the efficacy of DOM or must be considered prior to it's broader application. Through this effort, 19 "bridge" and 10 "barrier" categories were determined. For both the bridge and barrier code categories, three larger "umbrella" groups were established: species, social, and data. A table of the full list of codes/themes are depicted here. Please note that this is a draft of codes, and thus we are happy to modify the terminology throughout the study review process.

@@

```{r}
#| echo: false
#| message: false
#| warning: false

bri_codes <- code_freq %>% 
  filter(bridge_barr == "Bridge") %>% 
  subset(select = -c(total, bridge_barr))

colnames(bri_codes) <- c("Code category", "Code label")

bri_codes %>%
  kbl() %>%
  kable_styling()
```

@@

```{r}
#| echo: false
#| message: false
#| warning: false

bar_codes <- code_freq %>% 
  filter(bridge_barr == "Barrier") %>% 
  subset(select = -c(total, bridge_barr))

colnames(bar_codes) <- c("Code category", "Code label")

bar_codes %>%
  kbl() %>%
  kable_styling()
```
:::

## Visualizing the bridges and barrier codes from the literature

@@

### Code appearance frequencies across articles (n = 109)

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-align: center

codes_vars2 = codes_vars
colnames(codes_vars2) <- gsub(x = colnames(codes_vars2), pattern = "\\.", replacement = " ")

  #calculate frequency of bridges vs. barriers being mentioned in articles -- need to add biotic and abiotic indicators cols
code_freq_new <- codes_vars2 %>%
  mutate(br_pos = as.numeric(if_any(33:48, ~. == 1)), 
         ba_pos = as.numeric(if_any(49:58, ~. == 1)), 
         br_dat = as.numeric(if_any(c("Management scale aligns with system needs", "Target species data availability", "Access to advanced technology and techniques", "Data transparency and accessibility", "Abiotic data availability", "Resource use data availability"), ~. == 1)), 
         br_eco = as.numeric(if_any(c("Expected climate change induced range shifts", "Distinct habitat preferences among species in system", "Highly mobile target species", "Target species is a prey specialist", "Life history information available for target species", "Fine scale management appropriate for target species"), ~. == 1)), 
         br_soc = as.numeric(if_any(c("Resource availability", "Pre existing regulation and management institutions", "Adoption incentives for end users", "Feasibility", "Co management and communication"), ~. == 1)), 
         ba_dat = as.numeric(if_any(c("Implementation costs", "Gaps and inaccuracy", "Management scales do not align with system needs", "Model abuse", "Climate change adding uncertainty to target species habitat use"), ~. == 1)), 
         ba_eco = as.numeric(if_any(c("Need for multispecies management"), ~. == 1)), 
         ba_soc = as.numeric(if_any(c("Indirect socioeconomic consequences", "Lack of political administration permanence", "Poor communication and outreach", "Poor program implementation"), ~. == 1)))

brba_sum <- code_freq_new %>%
  summarise(n_br = sum(br_pos), freq_br = n_br/109,
            n_ba = sum(ba_pos), freq_ba = n_ba/109,
            n_br_e = sum(br_eco), freq_br_e = n_br_e/109,
            n_br_d = sum(br_dat), freq_br_d = n_br_d/109,
            n_br_s = sum(br_soc), freq_br_s = n_br_s/109,
            n_ba_e = sum(ba_eco), freq_ba_e = n_ba_e/109,
            n_ba_d = sum(ba_dat), freq_ba_d = n_ba_d/109,
            n_ba_s = sum(ba_soc), freq_ba_s = n_ba_s/109) %>%
  gather(br_ba, value, c(2,4,6, 8, 10, 12, 14, 16)) %>%
  subset(select = c(br_ba, value)) %>%
  mutate(percent_label = scales::percent(value, accuracy = 1), 
         percent = if_else(br_ba == "freq_ba" | br_ba == "freq_ba_e" | br_ba == "freq_ba_d" | br_ba == "freq_ba_s", -value, value),
         code_level = if_else(br_ba %in% c("freq_ba", "freq_br"), "Overall bridge or barrier", 
                              if_else(br_ba %in% c("freq_ba_e", "freq_br_e"), "Ecological bridge or barrier", 
                              if_else(br_ba %in% c("freq_ba_d", "freq_br_d"), "Data bridge or barrier",
                              if_else(br_ba %in% c("freq_ba_s", "freq_br_s"), "Social bridge or barrier", NA)))),
         code_brba = if_else(br_ba %in% c("freq_br", "freq_br_e", "freq_br_s", "freq_br_d"), "bridge", "barrier"),
         alpha_brba = if_else(br_ba %in% c("freq_br_e", "freq_br_s", "freq_br_d", "freq_ba_e", "freq_ba_s", "freq_ba_d"), "alpha", "no_alpha")) %>%
  group_by(code_level)

brba_sum$code_level <- factor(brba_sum$code_level, levels = c("Social bridge or barrier", "Ecological bridge or barrier", "Data bridge or barrier", "Overall bridge or barrier"))

ggplot(brba_sum, aes(x = code_level, 
             y = percent,
             fill = code_brba)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = percent_label),
            position = position_stack(vjust = 0.5),
            color = "black", 
            size = 5) +
  ylim(-0.6,1)+
  coord_flip() +
  scale_x_discrete() +
  scale_fill_manual(values = c("lightskyblue3", "goldenrod"), labels = c("Barrier frequency", "Bridge frequency")) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x =element_blank(),
        axis.text.y = element_text(size = 20, color = "black"),
        panel.grid = element_blank(),
        legend.position = c(-0.21, 0.99),
        legend.text = element_text(size = 16, color = "black"),
        strip.text.y.right = element_text(angle = 0, size = 16)) +
  geom_hline(yintercept = 0.0, linewidth = 1) + 
  labs(fill = "")+ 
  guides(fill = guide_legend(nrow = 1))
```
